---
title: "Merging/Cleaning Data Final Project"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# CRITERIA
- The submitted data set is tidy.
- The Github repo contains the required scripts.
- GitHub contains a code book that modifies and updates the available 
 codebooks with the data to indicate all the variables and summaries calculated, 
- along with units, and any other relevant information.
- The README that explains the analysis files is clear and understandable.
- The work submitted for this project is the work of the student who submitted it.


# Instructions As Code
*You should create one R script called run_analysis.R that does the following.*

```{r include=FALSE}
library(dplyr)
library(data.table)
```

*1. Merges the training and the test sets to create one data set.*

```{r}
X_train <- read.table("data/train/X_train.txt")
X_test <- read.table("data/test/X_test.txt")

y_train <- read.table("data/train/y_train.txt")
y_test <- read.table("data/test/y_test.txt")

subject_train <- read.table("data/train/subject_train.txt")
subject_test <- read.table("data/test/subject_test.txt")

X_combined <- bind_rows(X_train, X_test)
y_combined <- bind_rows(y_train, y_test)
subject_combined <- bind_rows(subject_train, subject_test)
```

*2. Extracts only the measurements on the mean and standard deviation for each measurement.*

First lets give them names so we can slice and dice.
Note that some of the names are invalid from an R perspective, so lets clean them up
```{r}
names_list <- read.csv("data/features.txt", sep=" ", header = FALSE)
colnames <- make.names(names=as.character(names_list[,2]), unique=TRUE, allow_ = TRUE)
setnames(X_combined, colnames)
```

Now lets extract
```{r}
X_mean_and_std <- X_combined %>% select(contains("mean"), contains("std"))
```

*3. Uses descriptive activity names to name the activities in the data set*
```{r}
activity_labels_list <- read.csv("data/activity_labels.txt", sep=" ", header = FALSE)
```

Marry in the y values for adding the activity names
```{r}
Xy_combined <- bind_cols(X_mean_and_std, y_combined)
Xy_combined <- Xy_combined %>% 
                  rename(Activity.Index=V1) %>% 
                  inner_join(activity_labels_list, by=c("Activity.Index" = "V1"))
```

Add in the subject ids
```{r}
Xy_combined <- bind_cols(Xy_combined, subject_combined)
Xy_combined <- Xy_combined %>% 
                rename(Subject.Id=V1, Activity.Name=V2) %>%
                select(-Activity.Index) #Remove activity index
str(Xy_combined)
```

*4. Appropriately labels the data set with descriptive variable names.*

```{r}
names(Xy_combined) <- sub("^f", "Frequency", names(Xy_combined))
names(Xy_combined) <- sub("^t", "Time", names(Xy_combined))
names(Xy_combined) <- sub("Acc", "Acceleration", names(Xy_combined))
names(Xy_combined) <- sub("Gyro", "Gyroscope", names(Xy_combined))
names(Xy_combined) <- sub("angle.t", "Angle.Time", names(Xy_combined))
names(Xy_combined) <- sub("angle", "Angle", names(Xy_combined))
names(Xy_combined) <- sub("BodyBody", "Body", names(Xy_combined))
names(Xy_combined)
```

*5. From the data set in step 4, creates a second, independent tidy data set with the 
   average of each variable for each activity and each subject.*
   
Assuming this is to be something like

subject activity  avg-tBodyAcc.mean...X tBodyAcc.mean...Y ...etc.

1       walking   0.3535  -0.131

where subject, activity are labels and measurements are variables
In other words, average each column, grouped by subject id, then activity

```{r}
tidy_averages <-aggregate(. ~Subject.Id + Activity.Name, Xy_combined, mean)
write.table(tidy_averages, file = "averages_per_subject_per_activity.txt", row.names=FALSE)
```